---
title: "Final Project - Partials"
author: "Donald Stephens (ds3478)"
date: "4/2/2022"
always_allow_html: yes
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(htmltools)
library(ggplot2)
library(ggthemes)
library(DT)
library(leaflet)
library(leaflet.extras)
library(sf)
#install.packages('rgdal')
library(rgdal) # for reading shape files into R
library(rgeos)
library(maps)
library(crosstalk)
library(vembedr)
```
## Formula 1 Racing

### Background Info on Formula 1

```{r, echo=FALSE, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE, cache = FALSE}
embed_url("https://www.youtube.com/watch?v=wtJPe1ksS6E¬")

#embed_url("https://www.youtube.com/watch?v=Z_LmiQlT9oE")
```

<hr/>

## Cleaned Data and Statistics

### Circuits

<p>
<span style="color:#00008B;font-size:12pt">
An interactive datatable of circuits across the globe. The original orientation is ordered by country and location. Some special additional features added include:
  <ol style="color:#00008B;font-size:12pt">
    <li>An <u>export button</u> that allows users to download data in CSV format preserving row order in the grid.</li>
    <li>A column with a <u>clickable link</u> that opens a new browser with information specific to that circuit.</li>
  </ol>
</span>
</p>
<br/>

```{r, echo=FALSE, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE, cache = FALSE}
circuits_type_df <- read_csv("../raw_data/circuit_type.csv")
circuits_type_df <- circuits_type_df %>%
  mutate(circuit_type =
    case_when(
      street_circuit == 1 ~ "street",
      road_circuit == 1 ~ "road",
      race_circuit == 1 ~ "race"
    )
  ) %>%
  dplyr::select(circuitId, circuit_type) %>%
  arrange(circuitId)

#View(circuits_type_df)
```

<p>
<span style="color:#00008B;font-size:12pt">
</span>
</p>

```{r, echo=FALSE, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE, cache = FALSE}
circuits_df <- read_csv("../raw_data/circuits.csv")
circuits_df <- circuits_df %>%
  mutate(information = paste("<a target='_blank' href='", url,"'>link</a>")) %>%
  dplyr::select(circuitId, name, location, country, lat, lng, alt, information) %>%
  arrange(country, location, name)

circuits_df$country <-
  ifelse(circuits_df$country == "America", "United States", circuits_df$country)

circuits_df$country <-
  ifelse(circuits_df$country == "USA", "United States", circuits_df$country)

circuits_df <- circuits_df %>%
  left_join(circuits_type_df, by = "circuitId") %>%
  dplyr::select(
    circuitId, name, location, country,
    circuit_type, lat, lng, alt, information
  ) %>%
  arrange(country, location, name)

#View(circuits_df)

##circuits_us_df <- circuits_df %>%
##  filter(country == "United States")

##View(circuits_us_df)

datatable(circuits_df,
  class = 'cell-border stripe',
  #colnames = c("Name", "Location", "Country"),
  caption = 'Table 1: Circuits',
  extensions = 'Buttons',
  options = list(
    pageLength = 5,
    columnDefs = list(list(visible=FALSE, targets=c("circuitId", "lat", "lng", "alt"))),
    dom = 'Bfrtrip',
    buttons = c('csv') # 'copy', 
  ),
  escape = FALSE
)
```

```{r, echo=FALSE, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE, cache = FALSE}
circuits_country_group_df <- circuits_df %>%
  group_by(country) %>% # circuit_type
  arrange(country)  %>% # circuit_type
  dplyr::summarize(num_circuits = n())
#View(circuits_country_group_df)
```

<br/><br/>
<p>
<span style="color:#00008B;font-size:12pt">
A distribution of circuits per country. Provides insight into countries with the most and least number of circuits. United States has the most number of circuits in the world at 12 followed by France at 7.
</span>
</p>

```{r, echo=FALSE, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE, cache = FALSE}
theme_set(theme_bw(16))

ggplot(data = circuits_country_group_df) + 
  geom_bar(
    mapping = aes(x = factor(country), y = num_circuits, fill = country), # circuit_type
    stat = "identity",
    position = "dodge",
    show.legend = FALSE #TRUE
  ) +
  scale_y_continuous(labels = scales::comma) + 
  ggtitle("Circuits by Country") +
  xlab("Country") +
  ylab("Number of Circuits") +
  #scale_fill_manual("legend", values = c("failed" = "dark red", "successful" = "dark green")) +
  theme(
    plot.title=element_text(hjust = 0.5),
    axis.text.x=element_text(angle = 90, vjust = 0.5),
    #legend.title = element_blank()
  ) #+
  #coord_flip()
```

#### Circuit Locations

<p>
<span style="color:#00008B;font-size:12pt">
For interactivity, added a zoomable geospatial map plotting circuits (with a custom icon) for a country selected by a dropdown (defaulting to all countries). Clicking on an icon displays some information about the circuit. The dropdown allows:
<ul style="color:#00008B;font-size:12pt">
  <li>multi-country selection</li>
  <li>manual typing for filtering and selection</li>
</ul>
</span>
</p>

```{r, echo=FALSE, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE}
# #racing_circuits <- geojsonio::geojson_read('f1-circuits.geojson',
# #  what = "sp")
# ##class(racing_circuits)
# ##names(racing_circuits)
# 
# #us_circuits <- geojsonio::geojson_read('circuits/us-2012.geojson',
# #  what = "sp")
# ##class(us_circuits)
# ##names(us_circuits)
# 
# avg_lat <- mean(circuits_us_df$lat) #circuits_df$lat)
# avg_long <- mean(circuits_us_df$lng) #circuits_df$lng)
# 
# leaflet() %>%
#   setView(lng = avg_long, lat = avg_lat, zoom = 4) %>%
#   addProviderTiles(providers$CartoDB.Voyager) %>% 
#   #addTiles() %>%
#   #addCircles(data = circuits_df,
#   #  lat = ~lat, lng = ~lng,
#   #  weight = 10, radius = 7, color = "blue") %>%
#   addMarkers(data = circuits_us_df, #circuits_df,
#     lat = ~lat, lng = ~lng,
#     icon = list(
#       iconUrl = 'track-icon-19.png',
#       iconSize = c(25, 25)
#     ),
#     popup = ~paste("<b>", name, "</b><br/><i>Type</i>: ", circuit_type, "<br/><i>Location</i>: ", location, "<br/><i>GPS</i>: (", lat, ",", lng, ",", alt, ")")) #%>%
#   #addPolygons(data = us_circuits, weight = 1, color = "light gray")
#   #addGeoJSON(racing_circuits) #us_circuits)
```

```{r, echo=FALSE, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE}
# https://stackoverflow.com/questions/47597012/dropdown-filter-in-leaflet-for-r
# https://www.rdocumentation.org/packages/crosstalk/versions/1.2.0/topics/bscols
# https://www.rdocumentation.org/packages/crosstalk/versions/1.2.0/topics/filter_select

country_sd <- SharedData$new(circuits_df)

map <- leaflet() %>%
  addProviderTiles(providers$CartoDB.Voyager) %>% 
  addMarkers(data = country_sd,
    lat = ~lat, lng = ~lng,
    icon = list(
      iconUrl = 'track-icon-19.png',
      iconSize = c(25, 25)
    ),
    popup = ~paste("<b>", name, "</b><br/><i>Type</i>: ", circuit_type, "<br/><i>Location</i>: ", location, ", ", country, "<br/><i>GPS</i>: (", lat, ",", lng, ",", alt, ")", sep = ""))

bscols(map)

bscols(
  filter_select("CountrySelector", "Country Circuit Selector:", country_sd, ~country)
)
```

<hr />

### Constructors

```{r, echo=FALSE, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE, cache = FALSE}
#constructor_results_df <- read_csv("../raw_data/constructor_results.csv")
#View(constructor_results_df)
```

```{r, echo=FALSE, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE, cache = FALSE}
constructors_df <- read_csv("../raw_data/constructors.csv")
constructors_df <- constructors_df %>%
  mutate(total_points = 0, point_2019 = 0, points_2021 = 0) %>%
  mutate(information = paste("<a target='_blank' href='", url,"'>link</a>")) %>%
  dplyr::select(
    constructorId, name, nationality,
    #total_points, point_2019, points_2021,
    information
  ) %>%
  arrange(nationality, name)
#View(constructors_df)

datatable(constructors_df,
  class = 'cell-border stripe',
  #colnames = c("name", "nationality"),
  caption = 'Table 2: Constructors',
  options = list(
    pageLength = 5,
    columnDefs = list(list(visible=FALSE, targets=c("constructorId")))
  ),
  escape = FALSE
)
```

```{r, echo=FALSE, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE, cache = FALSE}
circuits_country_group_df <- circuits_df %>%
  group_by(country) %>%
  arrange(country)  %>%
  dplyr::summarize(num_circuits = n())
#View(circuits_country_group_df)
```

<!-- <hr /> -->

<!-- ### Races -->

```{r, echo=FALSE, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE, cache = FALSE}
races_df <- read_csv("../raw_data/races.csv")
races_df <- races_df %>%
  filter(year >= 2019, year <= 2021, year != 2020) %>%
  dplyr::select(raceId, name, year, date, round, time, circuitId) %>%
  arrange(desc(year), desc(date), desc(round), time)
#View(races_df)

races_circuits_df <- races_df %>%
  left_join(circuits_df, by = "circuitId") %>%
  select(raceId, name.x, year, date, round, time, circuitId, name.y, location, country) %>%
  rename(race_name = name.x, circuit_name = name.y, circuit_location = location, circuit_country = country)
#View(races_circuits_df)

# datatable(races_circuits_df,
#   class = 'cell-border stripe',
#   #colnames = c("name", "nationality"),
#   caption = 'Races',
#   options = list(
#     pageLength = 5,
#     columnDefs = list(list(visible=FALSE, targets=c("raceId", "circuitId")))
#   )
# )
```

<hr />

### Drivers

```{r, echo=FALSE, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE, cache = FALSE}
drivers_df <- read_csv("../raw_data/drivers.csv")

drivers_df$number <-
  ifelse(drivers_df$number == "\\N", "", drivers_df$number)

drivers_df$code <-
  ifelse(drivers_df$code == "\\N", "", drivers_df$code)

drivers_df <- drivers_df %>%
  mutate(information = paste("<a target='_blank' href='", url,"'>link</a>")) %>%
  mutate(driver_number = number, driver_code = code) %>%
  mutate(firstname = forename, lastname = surname, birth_date = dob) %>%
  dplyr::select(driverId, firstname, lastname, birth_date,
    driver_number, driver_code, nationality, information
  ) %>%
  arrange(nationality, lastname, firstname, birth_date, driver_number)

#View(drivers_df)

datatable(drivers_df,
  class = 'cell-border stripe',
  #colnames = c("name", "nationality"),
  caption = 'Table 3: Drivers',
  options = list(
    pageLength = 5,
    columnDefs = list(list(visible=FALSE, targets=c("driverId")))
  ),
  escape = FALSE
)
```

<!-- #### Driver Standings -->

```{r, echo=FALSE, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE, cache = FALSE}
# # May need Server-side processing
# #  You may consider server-side processing: https://rstudio.github.io/DT/server.html
# 
# driver_standings_df <- read_csv("../raw_data/driver_standings.csv")
# driver_standings_df <- driver_standings_df %>%
#   arrange(desc(raceId), position)
# #View(driver_standings_df)
# 
# driver_standings_race_df <- driver_standings_df %>%
#   left_join(drivers_df, by = "driverId") %>%
#   left_join(races_df, by = "raceId") #%>%
# #  #select(driverStandingsId, raceId, driverId, name.z, points, poisition, wins)
# 
# datatable(driver_standings_race_df,
#   class = 'cell-border stripe',
#   #colnames = c("name", "nationality"),
#   caption = 'Driver Standings',
#   options = list(
#     pageLength = 5#,
#     #columnDefs = list(list(visible=FALSE, targets=c("driverStandingsId, raceId")))
#   )
# )
```

<!-- <hr /> -->

<!-- ### Results -->

```{r, echo=FALSE, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE, cache = FALSE}
# results_df <- read_csv("../raw_data/results.csv")
# results_df <- results_df %>%
#   dplyr::select(resultId, raceId, driverId, constructorId, rank,
#     positionOrder, points, laps, time, statusId,
#     fastestLap, fastestLapTime, fastestLapSpeed
#   )
# View(results_df)
# 
# datatable(results_df,
#   class = 'cell-border stripe',
#   #colnames = c("name", "nationality"),
#   caption = 'Results',
#   options = list(pageLength = 5)
# )
```

```{r, echo=FALSE, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE, cache = FALSE}
# results_merged_df <- results_df %>%
#  left_join(races_df, by = "raceId") %>%
#  rename(race_name = name)
# View(results_merged_df)
```