---
title: "f1_exploratory"
author: "Christa Chiao"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())
```

```{r}
# broken thing to call files but just run the `load_files_code.R` script that is
# in the scoping folder.

library(dplyr)
library(stringr)

setwd("C:/Users/chris/Dropbox/2021_QMSS/Spring 2022/GR5063 Data Visualization/group_f_f1/raw_data")

# if the files are in a different location than your working directory, change
# the file path in the list.files function
file_names <- list.files()

# read all 13 csvs into a list called "all_csvs"
all_csvs <- lapply(file_names, read.csv)

# Set the name of each list element to its respective file name.
names(all_csvs) <- gsub(".csv","",
                      list.files(full.names = FALSE),
                      fixed = TRUE)

# break list out into separate dfs
list2env(all_csvs, envir=.GlobalEnv)

```

## Sankey Diagram

### Individual's Sankey Diagram
```{r}
library(tidyr)
library(curl)
# install.packages("networkD3")
# Let's get data into the format we need. We need two dfs, one of nodes that has
# all of the options, which is 0-20, which is the positions you can start in and 
# the positions you can end in. and of links, which gives you the start and end
# positions, the count of times a driver has ended up in those positions, and
# the driver name

# shoudl also do this for constructor, maybe let's do that one first.



links_driver_2019 <- 
  results %>%
  left_join(races, by = "raceId") %>%
  filter(year == 2019) %>%
  left_join(drivers, by = "driverId") %>%
  unite("full_name", forename:surname, sep= " ", remove = FALSE) %>%
  select(raceId, driverId, full_name, positionOrder) %>%
  rename(race_position = positionOrder) %>%
  left_join(qualifying, by = c("raceId" = "raceId", 
                               "driverId" = "driverId"
                               )) %>%
  select(raceId, full_name, race_position, position) %>%
  rename(quali_position = position) %>%
  mutate(quali_position = as.integer(quali_position)) %>%
  mutate(race_position = if_else(race_position == "\\N", 0, as.numeric(race_position)),
         quali_position = if_else(is.na(quali_position), 0, as.numeric(quali_position))) %>%
  group_by(full_name, race_position, quali_position) %>%
  summarise(n = n()) %>%
  rename(target = race_position,
         source = quali_position,
         value = n) %>%
  mutate(source = source,
         target =  target) %>%
    mutate(target = as.integer(target),
         source = as.integer(source))

links_driver_2021 <- 
  results %>%
  left_join(races, by = "raceId") %>%
  filter(year == 2021) %>%
  left_join(drivers, by = "driverId") %>%
  unite("full_name", forename:surname, sep= " ", remove = FALSE) %>%
  select(raceId, driverId, full_name, positionOrder) %>%
  rename(race_position = positionOrder) %>%
  left_join(qualifying, by = c("raceId" = "raceId", 
                               "driverId" = "driverId"
                               )) %>%
  select(raceId, full_name, race_position, position) %>%
  rename(quali_position = position) %>%
  mutate(quali_position = as.integer(quali_position)) %>%
  mutate(race_position = if_else(race_position == "\\N", 0, as.numeric(race_position)),
         quali_position = if_else(is.na(quali_position), 0, as.numeric(quali_position))) %>%
  group_by(full_name, race_position, quali_position) %>%
  summarise(n = n()) %>%
  rename(target = race_position,
         source = quali_position,
         value = n) %>%
  mutate(source = source,
         target =  target) %>%
    mutate(target = as.integer(target),
         source = as.integer(source))

library(networkD3)
# save the widget
# library(htmlwidgets)
# saveWidget(p, file=paste0( getwd(), "/HtmlWidget/sankeyColor3.html"))

```

Max Verstappen Sankey Diagram 2019 and 2021
```{r}
links_driver_mv_19 <- links_driver_2019 %>% filter(full_name == "Max Verstappen")
links_driver_mv_19 <- as.data.frame(links_driver_mv_19)

nodes1 <- c("0","1","2","3","4","5","6","7","8","9","10",
"11","12","13","14","15","16","17","18","19","20")

nodes1 <- as.data.frame(nodes1)

my_color_1 <- 
  'd3.scaleOrdinal() .domain(["Max Verstappen"]) .range(["#1E41FF"])'

# 2019 Max Verstappen
mv_2019 <- sankeyNetwork(Links = links_driver_mv_19, Nodes = nodes1, Source = "source",
              Target = "target", Value = "value", 
              NodeID = "nodes1", colourScale=my_color_1, LinkGroup = "full_name",
              NodeGroup = NULL)

links_driver_mv_21 <- links_driver_2021 %>% filter(full_name == "Max Verstappen")
links_driver_mv_21 <- as.data.frame(links_driver_mv_21)

# 2021 Max Verstappen
mv_2021 <- sankeyNetwork(Links = links_driver_mv_21, Nodes = nodes1, Source = "source",
              Target = "target", Value = "value", 
              NodeID = "nodes1", colourScale=my_color_1, LinkGroup = "full_name",
              NodeGroup = NULL)

```


Lewis Hamilton Sankey Diagram 2019 and 2021

#### How often do the top two drivers convert their Qualifying positions to Wins?

One way to gauge how successful the top two drivers are is how often they are able to convert their pole positions into a win. Both Hamilton and Verstappen are exceptional drivers, but in 2019 did not qualify well, Hamilton only earning five pole positions and Max Verstappen earning two pole positions. Hamilton was able to convert three of the five pole positions he earned to a race victory, securing his other eight victories outside of pole position. Max Verstappen converted only one of his two pole positions into race victories, earning two race wins outside of poll position. 

This sankey diagram shows in the 2019 season, which final race position Lewis Hamilton and Max Verstappen end up in after starting in pole. The teal represents a Mercedes driver, Lewis Hamilton, while the darker blue represents a Red Bull driver, Max Verstappen. 

```{r, echo=FALSE, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE, cache = FALSE}
links_driver_lhmv_2019 <- 
  links_driver_2019 %>% 
  filter(full_name == "Lewis Hamilton" | full_name == "Max Verstappen") %>%
  filter(source == 1)
  

links_driver_lhmv_2019 <- as.data.frame(links_driver_lhmv_2019)


nodes19 <- c("0","1","2","3","4","5","6","7","8","9")

nodes19 <- as.data.frame(nodes19)




my_color <-
  'd3.scaleOrdinal() .domain(["Lewis Hamilton", "Max Verstappen"]) .range(["#6CD3BF", "#1E41FF"])'

sankeyNetwork(Links = links_driver_lhmv_2019, Nodes = nodes19, Source = "source",
              Target = "target", Value = "value",
              NodeID = "nodes19", fontSize= 25, nodeWidth = 20,
              colourScale=my_color, LinkGroup="full_name",
              NodeGroup=NULL)
```


In the 2021 season, Max Verstappen surged in his qualifying skills, earning ten pole positions to Lewis Hamilton's five pole positions. Max was able to convert seven of his pole positions to a victory, earning three race wins outside of pole position, while Lewis Hamilton converted three of his pole positions to a victory, earning five race wins outside of pole.

This sankey diagram shows  which final race position Lewis Hamilton and Max Verstappen end up in after starting in pole during the 2021 season. 

```{r, echo=FALSE, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE, cache = FALSE}

links_driver_lhmv_2021 <- 
  links_driver_2021 %>% 
  filter(full_name == "Lewis Hamilton" | full_name == "Max Verstappen")  %>%
  filter(source == 1)

links_driver_lhmv_2021 <- as.data.frame(links_driver_lhmv_2021)

nodes21 <- c("0","1","2","3","4","5")

nodes21 <- as.data.frame(nodes21)

my_color <-
  'd3.scaleOrdinal() .domain(["Lewis Hamilton", "Max Verstappen"]) .range(["#6CD3BF", "#1E41FF"])'

sankeyNetwork(Links = links_driver_lhmv_2021, Nodes = nodes21, Source = "source",
              Target = "target", Value = "value",
              NodeID = "nodes21", fontSize= 25, nodeWidth = 20, 
              colourScale=my_color, LinkGroup = "full_name",
              NodeGroup = NULL)


Both Max Verstappen and Lewis Hamilton Sankey Diagram
```




```{r}
links_driver_lhmv_2019 <- 
  links_driver_2019 %>% 
  filter(full_name == "Lewis Hamilton" | full_name == "Max Verstappen") %>%
  filter(source == 1)
  
nodes1 <- c("0","1","2","3","4","5","6","7","8","9","10",
"11","12","13","14","15","16","17","18","19","20")

nodes1 <- as.data.frame(nodes1)


links_driver_lhmv_2019 <- as.data.frame(links_driver_lhmv_2019)

my_color <-
  'd3.scaleOrdinal() .domain(["Lewis Hamilton", "Max Verstappen"]) .range(["#6CD3BF", "#1E41FF"])'

sankeyNetwork(Links = links_driver_lhmv_2019, Nodes = nodes1, Source = "source",
              Target = "target", Value = "value",
              NodeID = "nodes1", fontSize= 25, nodeWidth = 20, margin = list(left = 100),
              colourScale=my_color, LinkGroup="full_name",
              NodeGroup=NULL)
```


```{r}
links_driver_lhmv_2021 <- 
  links_driver_2021 %>% 
  filter(full_name == "Lewis Hamilton" | full_name == "Max Verstappen")  %>%
  filter(source == 1)

links_driver_lhmv_2021 <- as.data.frame(links_driver_lhmv_2021)

nodes21 <- c("0","1","2","3","4","5","6","7","8","9","10",
"11","12","13","14","15","16","17","18","19","20")

nodes21 <- as.data.frame(nodes21)

my_color <-
  'd3.scaleOrdinal() .domain(["Lewis Hamilton", "Max Verstappen"]) .range(["#6CD3BF", "#1E41FF"])'

sankeyNetwork(Links = links_driver_lhmv_2021, Nodes = nodes21, Source = "source",
              Target = "target", Value = "value",
              NodeID = "nodes21", fontSize= 15, nodeWidth = 20, margin = list(left = 100),
              colourScale=my_color, LinkGroup = "full_name",
              NodeGroup = NULL)
```


Constructor's Sankey Diagram - Helper Table
```{r}
links_const_2019 <- 
  results %>%
  left_join(races, by = "raceId") %>%
  filter(year == 2019) %>%
  left_join(constructors, by = "constructorId") %>%
  rename(const_name = name.y) %>%
  select(raceId, constructorId, const_name, position) %>%
  rename(race_position = position) %>%
  left_join(qualifying, by = c("raceId" = "raceId", 
                               "constructorId" = "constructorId"
                               )) %>%
  select(raceId, const_name, race_position, position) %>%
  rename(quali_position = position) %>%
  mutate(quali_position = as.integer(quali_position)) %>%
  mutate(race_position = if_else(race_position == "\\N", 0, as.numeric(race_position)),
         quali_position = if_else(is.na(quali_position), 0, as.numeric(quali_position))) %>%
  group_by(const_name, race_position, quali_position) %>%
  summarise(n = n()) %>%
  rename(target = race_position,
         source = quali_position,
         value = n) %>%
  mutate(source = source,
         target =  target) %>%
    mutate(target = as.integer(target),
         source = as.integer(source))

links_const_2021 <- 
  results %>%
  left_join(races, by = "raceId") %>%
  filter(year == 2021) %>%
  left_join(constructors, by = "constructorId") %>%
  rename(const_name = name.y) %>%
  select(raceId, constructorId, const_name, position) %>%
  rename(race_position = position) %>%
  left_join(qualifying, by = c("raceId" = "raceId", 
                               "constructorId" = "constructorId"
                               )) %>%
  select(raceId, const_name, race_position, position) %>%
  rename(quali_position = position) %>%
  mutate(quali_position = as.integer(quali_position)) %>%
  mutate(race_position = if_else(race_position == "\\N", 0, as.numeric(race_position)),
         quali_position = if_else(is.na(quali_position), 0, as.numeric(quali_position))) %>%
  group_by(const_name, race_position, quali_position) %>%
  summarise(n = n()) %>%
  rename(target = race_position,
         source = quali_position,
         value = n) %>%
  mutate(source = source,
         target =  target) %>%
    mutate(target = as.integer(target),
         source = as.integer(source))


library(networkD3)
```

Mercedes' Constructors Sankey Diagram 2019 and 2021
```{r}
links_const_merc19 <- links_const_2019 %>% filter(const_name == "Mercedes")
links_const_merc19 <- as.data.frame(links_const_merc19)

my_color_0 <- 
  'd3.scaleOrdinal() .domain(["Mercedes"]) .range(["#6CD3BF"])'

merc_2019 <- 
  sankeyNetwork(Links = links_const_merc19, Nodes = nodes1, Source = "source",
                Target = "target", Value = "value",
                NodeID = "nodes1", colourScale=my_color_0, LinkGroup = "const_name",
                NodeGroup = NULL)

links_const_merc21 <- links_const_2021 %>% filter(const_name == "Mercedes")
links_const_merc21 <- as.data.frame(links_const_merc21)

merc_2021 <- 
  sankeyNetwork(Links = links_const_merc21, Nodes = nodes1, Source = "source",
                Target = "target", Value = "value", 
                NodeID = "nodes1", colourScale=my_color_0, LinkGroup = "const_name",
                NodeGroup = NULL)
```

Red Bull Constructor Sankey Diagram 2019 and 2021
```{r}
links_const_rb19 <- links_const_2019 %>% filter(const_name == "Red Bull")
links_const_rb19 <- as.data.frame(links_const_rb19)

my_color_1 <- 
  'd3.scaleOrdinal() .domain(["Red Bull"]) .range(["#1E5BC6"])'

rb_2019 <- 
  sankeyNetwork(Links = links_const_rb19, Nodes = nodes1, Source = "source",
                Target = "target", Value = "value", 
                NodeID = "nodes1", colourScale=my_color_1, LinkGroup = "const_name",
                NodeGroup = NULL)


links_const_rb21 <- links_const_2021 %>% filter(const_name == "Red Bull")
links_const_rb21 <- as.data.frame(links_const_rb21)


rb_2021 <- 
  sankeyNetwork(Links = links_const_rb21, Nodes = nodes1, Source = "source",
                Target = "target", Value = "value", 
                NodeID = "nodes1", colourScale=my_color_1, LinkGroup = "const_name",
                NodeGroup = NULL)
```

All Constructor Sankey Diagram
```{r}
my_color <- 
  'd3.scaleOrdinal() .domain(["Alfa Romeo", "Alpha Tauri", "Alpine F1 Team", "Aston Martin", "Ferrari", "Haas F1 Team", "McLaren", "Mercedes", "Racing Point", "Red Bull", "Renault", "Toro Rosso", "Williams"]) .range(["#B12039", "#4E7C9B", "#2293D1", "#2D826D", "#ED1C24", "#B6BABD", "#F58020", "#6CD3BF", "#F596C8", "#1E41FF", "#FFF500", "#469BFF", "#FFFFFF"])'

const_2019 <- 
  sankeyNetwork(Links = links_const_2019, Nodes = nodes1, Source = "source",
                Target = "target", Value = "value", 
                NodeID = "nodes1", fontSize= 15, nodeWidth = 20, margin = list(left = 100),
                colourScale=my_color, LinkGroup = "const_name",
                NodeGroup = NULL)


const_2021 <- 
  sankeyNetwork(Links = links_const_2021, Nodes = nodes1, Source = "source",
                Target = "target", Value = "value",
                NodeID = "nodes1", fontSize= 15, nodeWidth = 20, margin = list(left = 100),
                colourScale=my_color, LinkGroup = "const_name",
                NodeGroup = NULL)

const_2019
```



### Helper Tables
```{r}
library(dplyr)

circuits_joined <- circuits %>%
                  left_join(circuit_type) %>%
                  rename(circuit_name = name)

circuits_by_year <- races %>%
  select(raceId, year, circuitId, name) %>%
  rename(race_name = name) %>%
  left_join(circuits_joined, by = c("circuitId")) %>%
  mutate(circuit_type = 
           case_when(street_circuit == 1 ~ "street",
                     road_circuit == 1 ~ "street",
                     race_circuit == 1 ~ "race",
                     )
         ) %>%
  select(year, circuitId, circuit_name, circuit_type)

skinny_races_circuits <- races %>%
  rename(race_name = name) %>%
  select(raceId, circuitId, race_name, year)

xover <- 
  results %>%
  left_join(races, by = ("raceId")) %>%
  filter(year == 2019 | year == 2021) %>%
  select(driverId, constructorId, year) %>%
  left_join(drivers, by = ("driverId")) %>%
  left_join(constructors, by = ("constructorId")) %>%
  mutate(first_initial = paste0(substr(forename, 1, 1), ".")) %>%
  mutate(full_name = paste(first_initial, surname)) %>%
  rename(constructor_name = name) %>%
  select(driverId, full_name, constructorId, constructor_name, year) %>%
  group_by(driverId, full_name, constructorId, constructor_name, year) %>% 
  slice(1)

xover <- 
  xover %>%
  mutate(full_name = if_else(full_name == "K. Räikkönen", "Kimi Raikkonen", full_name),
         full_name = if_else(full_name == "N. Hülkenberg", "Nico Hulkenberg", full_name),
         full_name = if_else(full_name == "S. Pérez", "Sergio Perez", full_name))

```


This plot_ly graph shows how many times drivers have gotten the "Fastest Lap" designation. The "Fastest Lap" designation gives the driver and the constructor extra points and monetary bonuses. 
```{r}
lap_times_by_driver <- 
  lap_times %>%
  left_join(skinny_races_circuits, by = ("raceId")) %>%
  filter(year == 2019 | year == 2021) %>%
  left_join(xover, by = c("driverId" = "driverId",
                          "year" = "year")) %>%
  group_by(year, raceId, driverId, full_name, constructor_name) %>%
  summarise(min(milliseconds)) %>%
  rename(fastestLap_ms = "min(milliseconds)") %>%
  mutate(fastestLap_s = fastestLap_ms/1000) %>%
  group_by(year, raceId) %>%
  mutate(rank = rank(fastestLap_s)) %>%
  filter(rank == 1) %>%
  group_by(year, full_name, constructor_name) %>%
  summarise(count_of_fastest_lap = n()) %>%
  mutate(fast_laps_2019 = if_else(year == "2019", count_of_fastest_lap, as.integer(0)),
         fast_laps_2021 = if_else(year == "2021", count_of_fastest_lap, as.integer(0))) %>%
  group_by(full_name, constructor_name) %>%
  summarise(year_2019 = sum(fast_laps_2019),
            year_2021 = sum(fast_laps_2021))

library(ggplot2)
library(plotly)

lap_times_by_driver %>%
  plot_ly(type = "bar",
          x = ~year_2019, 
          y = ~full_name,
          marker = list(color = 
                          c("#F4767B", # Charles LeClerc
                            "#F9B279", # Ricciardo
                            "#D3D5D7", # K Mag (Haas)
                            "#A6E4D8", # Lewis Hamilton; Mercedes
                            "#F58020", # Lando
                            "#789CDC", # Max Verstappen; Red Bull
                            "#4E7C9B", # Gasley Alpha Tauri
                            "#F4767B", # Vettel Ferrari 2019; Aston Martin 2021
                            "#F9C0DE", # Perez for Racing Point
                            "#A6E4D8"  # Bottas
                            )), 
          text = ~constructor_name,
          hovertemplate = paste('<b>Driver</b>: %{y}<br>',
                                '<b>Team</b>: %{text}<br>',
                                '<b>Year</b>: 2019<br>', 
                                 'Number of Fastest Laps</b>: %{x}<br><extra></extra>'),
          name = '2019 Season') %>%
  add_trace(type = "bar",
            x = ~year_2021, 
            y = ~full_name,
            marker = list(color = 
                            c("#ED1C24", # LeClerc
                              "#F58020", # Ricciardo Renault 2019, 2021 at McL
                              "#B6BABD", # Haas - Magnussen
                              "#6CD3BF", # Lewis Hamilton
                              "#F58020", # Norris, McLaren
                              "#1E41FF", # Max Verstappen; Red Bull
                              "#4E7C9B", # Gasly Toro Rosso 2019/Alpha Tauri 2021
                              "#2D826D", # Vettel but for Aston martin
                              "#1E5BC6", # Perez, Racing Point 2019, RB 2021
                              "#6CD3BF" # Bottas, Mercedes

                                 )),
          text = ~constructor_name,
          hovertemplate = paste('<b>Driver</b>: %{y}<br>',
                                '<b>Team</b>: %{text}<br>',
                                '<b>Year</b>: 2021<br>',
                                'Number of Fastest Laps</b>: %{x}<br><extra></extra>'),
          name = '2021 Season') %>%
  layout(title = 'Fastest Laps Designation', 
         xaxis = list(title = 'Count of Fastest Laps'),
         yaxis = list(title = ''),
         barmode = 'group',
         showlegend = FALSE)

```

```{r}
lap_times_by_driver_2019 <- 
  lap_times %>%
  left_join(skinny_races_circuits, by = ("raceId")) %>%
  filter(year == 2019) %>%
  left_join(xover, by = c("driverId" = "driverId",
                          "year" = "year")) %>%
  group_by(year, raceId, driverId, full_name, constructor_name) %>%
  summarise(min(milliseconds)) %>%
  rename(fastestLap_ms = "min(milliseconds)") %>%
  mutate(fastestLap_s = fastestLap_ms/1000) %>%
  group_by(year, raceId) %>%
  mutate(rank = rank(fastestLap_s)) %>%
  filter(rank == 1) %>%
  group_by(full_name, constructor_name) %>%
  summarise(count_of_fastest_lap = n()) %>%
  arrange()

library(ggplot2)
library(plotly)

lap_times_by_driver_2019 %>%
  plot_ly(type = "bar",
          x = ~count_of_fastest_lap, 
          y = ~full_name,
          marker = list(color =c("#DC0000", # Charles LeClerc
                                 "#F0D787", # K Mag (Haas)
                                 "#00D2BE", # Lewis Hamilton; Mercedes
                                 "#1E41FF", # Max Verstappen; Red Bull
                                 "#DC0000", # Vettel Ferrari 2019; Aston Martin 2021
                                 "#00D2BE")),
          text = ~constructor_name,
          hovertemplate = paste('<b>Driver</b>: %{y}<br>',
                                 '<b>Team</b>: %{text}<br>', 
                                 'Number of Fastest Laps</b>: %{x}<br><extra></extra>')) %>%
  layout(title = 'Fastest Laps in 2019', xaxis = list(title = 'Count of Fastest Laps'), 
         yaxis = list(title = 'Driver'))

```


```{r}

lap_times_by_driver_2021 <- 
  lap_times %>%
  left_join(skinny_races_circuits, by = ("raceId")) %>%
  filter(year == 2021) %>%
  left_join(xover, by = c("driverId" = "driverId",
                          "year" = "year")) %>%
  group_by(year, raceId, driverId, full_name, constructor_name) %>%
  summarise(min(milliseconds)) %>%
  rename(fastestLap_ms = "min(milliseconds)") %>%
  mutate(fastestLap_s = fastestLap_ms/1000) %>%
  group_by(year, raceId) %>%
  mutate(rank = rank(fastestLap_s)) %>%
  filter(rank == 1) %>%
  group_by(full_name, constructor_name) %>%
  summarise(count_of_fastest_lap = n()) %>%
  arrange()

lap_times_by_driver_2021 %>%
  plot_ly(type = "bar",
          x = ~count_of_fastest_lap, 
          y = ~full_name,
          marker = list(color =c("#F58020", # Ricciardo Renault 2019, 2021 at McL
                                 "#F58020", # Norris, McLaren
                                 "#00D2BE", # Lewis Hamilton; Mercedes
                                 "#1E41FF", # Max Verstappen; Red Bull
                                 "#4E7C9B", # Gasly Toro Rosso 2019/Alpha Tauri 2021
                                 "#1E5BC6", # Perez, Racing Point 2019, RB 2021
                                 "#00D2BE" # Bottas, Mercedes
                                 )),
          text = ~constructor_name,
          hovertemplate = paste('<b>Driver</b>: %{y}<br>',
                                 '<b>Team</b>: %{text}<br>', 
                                 'Number of Fastest Laps</b>: %{x}<br><extra></extra>')) %>%
  layout(title = 'Fastest Laps in 2021', xaxis = list(title = 'Count of Fastest Laps'), 
         yaxis = list(title = 'Driver'))
```


Now I'm interested in seeing the average time per lap over the course of a race by circuit.
```{r}
lap_time_by_lap <- lap_times %>%
  drop_na(milliseconds) %>%
  group_by(raceId, lap, time) %>%
  summarise(mean(milliseconds)/1000) %>%
  rename(mean_lap_speed_ms = `mean(milliseconds)/1000`) %>%
  left_join(skinny_races_circuits, by = ("raceId")) %>%
  left_join(circuits_joined, by = ("circuitId")) %>%
  select(raceId, lap, mean_lap_speed_ms, time, circuitId, race_name, year,
          circuit_name) %>%
    filter(year == 2019 | year == 2021) %>%
  mutate_at(c("year"), as.factor)

library(RColorBrewer)

lap_time_by_lap %>%
  ggplot(aes(x = lap, y = mean_lap_speed_ms, color = year)) +
  geom_point() + ylim(0,300) +
  scale_fill_brewer(palette = "Spectral") +
  facet_wrap(vars(circuit_name))
```


```{r}
library(lubridate)
incidents_by_circuit <- 
  results %>%
  left_join(status2, by = ("statusId")) %>%
  left_join(races, by = ("raceId")) %>%
  rename(race_name = name) %>%
  filter(year == 2019 | year == 2021) %>%
  select(date, raceId, race_name, driverId, constructorId, statusId, status) %>%
  left_join(xover, by = ("constructorId")) %>%
  group_by(year, date, raceId, race_name, constructor_name, status) %>%
  filter(status == "Collision" | status == "Accident") %>%
  select(year, date, race_name, full_name, constructor_name, status)

incidents_by_circuit$date <- lubridate::ymd(incidents_by_circuit$date)
```

2019 and 2021 breakdown of incidents by constructor
```{r}
incidents_by_circuit %>%
  filter(year == 2019 | year == 2021) %>%
  mutate(col_sum_2019 = if_else(year == 2019, 1, 0),
         col_sum_2021 = if_else(year == 2021, 1, 0)) %>%
  group_by(constructor_name) %>%
  summarise(year_2019 = sum(col_sum_2019),
            year_2021 = sum(col_sum_2021)) %>%
  plot_ly(type = "bar", name = '2019',
          x = ~constructor_name, 
          y = ~year_2019,
          marker = list(color =  c("#D07988",
                                    "#94B0C3",
                                    "#7ABEE3",
                                    "#81B4A7",
                                    "#F4767B",
                                    "#D3D5D7",
                                    "#F9B279",
                                    "#A6E4D8",
                                    "#F9C0DE",
                                    "#789CDC",
                                    "#FFF966",
                                    "#90C3FF",
                                    "#87D8EA"
                                  )),
          hovertemplate = paste('<b>Team</b>: %{x}<br>',
                                '<b>Year</b>: 2019<br>',
                                 'Number of Accidents or Collisions</b>: %{y}<br>')) %>%
  add_trace(y = ~year_2021, name = '2021', 
            marker = list(color = c("#9B0000", # Alfa Romeo
                                   "#4E7C9B", # Alpha Tauri
                                   "#2293D1", # Alpine
                                   "#2D826D", # Aston Martin
                                   "#ED1C24", # Ferrari
                                   "#B6BABD", # Haas
                                   "#F58020", # McLaren
                                   "#00D2BE", # Mercedes
                                   "#F596C8", # Racing Point
                                   "#1E5BC6", # Red Bull
                                   "#FFF500", # Renault
                                   "#469BFF", # Toro Rosso
                                   "#37BEDD" # Williams
                                   )
                          ),
            hovertemplate = paste('<b>Team</b>: %{x}<br>',
                                '<b>Year</b>: 2021<br>',
                                 'Number of Accidents or Collisions</b>: %{y}<br>')) %>%
  layout(title = 'Incidents by Constructor',
         xaxis = list(title = '<br><br><br>Constructor'), 
         yaxis = list(title = 'Count of Accidents ors Collisions'),
         barmode = 'group')
```

2021 breakdown of incidents by constructor
```{r}
incidents_by_circuit %>%
  filter(year == 2019) %>%
  group_by(constructor_name) %>%
  summarise(n = n()) %>%
  plot_ly(type = "bar",
          x = ~constructor_name, 
          y = ~n,
          marker = list(color =  c("#B12039", # Alfas Romeo
                                   "#ED1C24", # Ferrari
                                   "#B6BABD", # Haas
                                   "#F58020", # McLaren
                                   "#6CD3BF", # Mercedes
                                   "#F596C8", # Racing Point
                                   "#1E5BC6", # Red Bull
                                   "#FFF500", # Renault
                                   "#1E5BC6", # Red Bull
                                   "#469BFF", # Toro Rosso
                                   "#37BEDD" # Williams
                                   
                                  )),
          # text = ~constructor_name,
          hovertemplate = paste('<b>Team</b>: %{x}<br>', 
                                 'Number of Accidents or Collisions</b>: %{y}<br>')) %>%
  layout(title = 'Incidents by Constructor in 2019',
         xaxis = list(title = 'Constructor'), 
         yaxis = list(title = 'Count of Accidents or Collisions'))


incidents_by_circuit %>%
  filter(year == 2021) %>%
  group_by(constructor_name) %>%
  summarise(n = n()) %>%
  plot_ly(type = "bar",
          x = ~constructor_name, 
          y = ~n,
          marker = list(color =  c("#B12039", # Alfa Romeo
                                   "#4E7C9B", # Alpha Tauri
                                   "#2293D1", # Alpine
                                   "#2D826D", # Aston Martin
                                   "#ED1C24", # Ferrari
                                   "#B6BABD", # Haas
                                   "#F58020", # McLaren
                                   "#6CD3BF", # Mercedes
                                   "#1E5BC6", # Red Bull
                                   "#37BEDD" # Williams
                                   
                                  )),
          # text = ~constructor_name,
          hovertemplate = paste('<b>Team</b>: %{x}<br>', 
                                 'Number of Accidents or Collisions</b>: %{y}<br>')) %>%
  layout(title = 'Incidents by Constructor in 2021',
         xaxis = list(title = 'Constructor'), 
         yaxis = list(title = 'Count of Accidents or Collisions'))

```
